---
title: "NFL Tree and Cap Spending"
author: "Kathryn Daugherty"
date: '2019-09-28'
css: []
description: ''
highlight: yes
scripts: []
slug: nfl-tree-and-cap-spending
tags: []
categories: []
---

Let's explore the package 'collapsableTree' using NFL data.  I foraged the internet to create a fairly simple csv file with columns Sport, Division, Region, Team, and Cap_Spend.  The values in Cap_Spend are the team's active (2020) active salary spends per <a href="http://overthecap.com/salary-cap-space/">OverTheCap.com</a>.

Let's get R set-up and then we can explore.

```{r}
# Set CRAN Mirror
options(repos = c(CRAN = "http://cran.rstudio.com"))

# Set time zone
options(tz="America/New_York")
```

We're going to use the collapsableTree library to create the visualization.  Documentation is available <a href="https://adeelk93.github.io/collapsibleTree/">here</a> and <a href="https://cran.r-project.org/web/packages/collapsibleTree/collapsibleTree.pdf">here</a>.

```{r}
# Load data set 
library(readxl)
Football_Caps <- read_excel("~/Desktop/Football_Caps.xls")
head(Football_Caps)

# Load needed libraries
library(collapsibleTree)
```

The basic tree is fairly straight-forward to create.

```{r}
collapsibleTree(Football_Caps,
            hierarchy = c("Division","Region","Team"),
            width = 500)
```

(If the above image does not show on your screen, click <a href="http://rpubs.com/Kathryn_Ann_22/533407">HERE</a> instead.)

Let's also show the salary cap information that we added.  We'll need the dplyr package to do this.  To get the cumulative spending for each branch of the hierarchy, we'll use a sum function.  Make sure you hover over each node and notice that the exact salary spending is listed.

```{r}
library(dplyr)

Football_Caps %>%
  group_by(Division, Region, Team) %>%
  summarize(`Salary Spending` = sum(Cap_Spend)) %>%
collapsibleTreeSummary(
  hierarchy = c("Division","Region","Team"),
  root = "Football_Caps",
  width = 500,
  attribute = "Salary Spending")
```

What if we just wanted variation in the colors.  I like blue, so we'll have each node be a shade of blue.

```{r}
divisionColors <- RColorBrewer::brewer.pal(length(unique(Football_Caps$Division)), "Blues")
# Regions will be a gradient that resets between divisions
regionColors <- Football_Caps %>%
  arrange(Division, Region) %>% 
  group_by(Division) %>%
  distinct(Region) %>%
  mutate(colors = colorspace::sequential_hcl(length(Region))[seq_along(Region)])
# Teams will also be a gradient that resets between divisions, but not Regions
teamColors <- Football_Caps %>%
  arrange(Division, Region) %>% 
  group_by(Division) %>%
  distinct(Team) %>%
  mutate(colors = colorspace::sequential_hcl(length(Team))[seq_along(Team)])

Football_Caps %>%
  arrange(Division, Region, Team) %>%
  collapsibleTree(
    hierarchy = c("Division", "Region", "Team"),
    root = "Football_Caps",
    width = 500,
    fill = c(divisionColors, regionColors$colors, teamColors$colors))
```

That's all for now...

[THE END]

