---
title: Hurricane Dorian in the News
author: Kathryn Daugherty
date: '2019-09-22'
slug: hurricane-dorian-in-the-news
categories: []
tags: []
css: []
description: ''
highlight: yes
scripts: []
---

The 'newsflash' package is something I played with back in 2017 during Hurricane Irma.  Having recently been under watch for Dorian, I'm revisiting it to visualize it's news coverage.

```{r}
# Set CRAN Mirror
options(repos = c(CRAN = "http://cran.rstudio.com"))

# Set time zone
options(tz="America/New_York")

# Newsflash documentation
# https://github.com/hrbrmstr/newsflash

# Install packages
# devtools::install_github("hrbrmstr/newsflash")
# install.packages("tidyverse")
# install.packages("ggalt")
# install.packages("brbrthemes")
# install.packages("anytime")

# Load libraries
library(newsflash)
library(tidyverse)
library(ggalt)
library(hrbrthemes)
library(anytime) # Get errors without this
```

Let's look at Chryons before we examine network TV.

```{r}
# Look at chryons (captioning)
ch <- read_chyrons("2019-09-02")
head(ch)
```

Within the news on Labor Day, how often was Dorian mentioned?  We'll create a column just for that purpose.

```{r}
# Examine where Dorian mentioned ("mention" column)
# Convert time to hourly chunks ("hour" column)
ch2 <- mutate(ch, 
  hour = lubridate::hour(ts),
  text = tolower(text),
  mention = grepl("dorian", text))
head(ch2)
```

Now, let's visualize this to see how it trended throughout the day.

```{r}
ch2 %>% filter(mention) %>% 
  count(hour, channel) %>% 
  ggplot(aes(hour, n)) +
  geom_segment(aes(xend=hour, yend=0), 
               color = "lightslategray", size=1) + 
  scale_x_continuous(name="Hour (GMT)", breaks=seq(0, 23, 6),
                   labels=sprintf("%02d:00", seq(0, 23, 6))) +
  scale_y_continuous(name="# Chyrons", limits=c(0,20)) +
  facet_wrap(~channel, scales="free") +
  labs(title="'Dorian' Mentions per Hour on 2019 Labor Day") 
```

Next, let's examine networks.  First, lets see a sample of what's even available.

```{r}
# How about networks?
ne <- list_networks(widget=FALSE)
head(ne,20)
```

Similar to our chryon analysis, let's plot mentions over time.  Let's look at news coverage across a three week period instead.

```{r}
# Query Dorian data
# Primary term = Dorian
# Secondary term = Hurricane
# Starting just before Labor Day weekend
# https://rdrr.io/github/hrbrmstr/newsflash/man/query_tv.html
dorian <- query_tv("Dorian", mode = "TimelineVol",
                   start_date = "2019-08-27", end_date = "2019-09-12")
head(dorian)
```

Let's plot this, breaking up the different networks into individual facets.

```{r}
# Visualize network coverage over 3 weeks
# Themes causing errors, so commenting out parts
# https://github.com/hrbrmstr/hrbrthemes
query_tv('Dorian', mode = "TimelineVol",
                start_date = "2019-08-27", end_date = "2019-09-12") %>% 
        arrange(date) %>% 
        ggplot(aes(date, value, group=network)) +
        ggalt::geom_xspline(aes(color=network)) +
        ggthemes::scale_color_tableau(name=NULL) +
        labs(x=NULL, y="Volume Metric", 
             title="'Dorian' Trends Across National Networks") +
        facet_wrap(~network) +
        # theme_ipsum_rc(grid="XY") +
        theme(legend.position="none") 
```

We see that coverage built up, was significant for a while, and then quickly dropped off.

Another way the 'newsflash' package allows us to examine this topic is via a word cloud.  When Dorian was mentioned, what else was said?

```{r}
# What are the top words associated with Dorian?
# Let's do a word cloud
wc <- query_tv('Dorian', mode = "WordCloud", 
               start_date = "2019-09-01", end_date = "2019-09-03")
head(wc)
```

Next, let's plot this!

```{r}
ggplot(wc, aes(x=1, y=1)) +
        ggrepel::geom_label_repel(aes(label=label, size=count), segment.colour="#00000000", segment.size=0) +
        scale_size_continuous(trans="sqrt") +
        labs(x=NULL, y=NULL) +
        theme_ipsum_rc(grid="") +
        theme(axis.text=element_blank()) +
        theme(legend.position="none") 
```

This isn't my favorite word cloud viz, but it gets the job done.

[THE END]