---
title: Starbucks iOs App Analysis
author: Kathryn Daugherty
date: '2019-09-16'
slug: starbucks-ios-app-analysis
categories: []
tags: []
css: []
description: ''
highlight: yes
scripts: []
---

Have you ever wanted to see more detail behind apps on your iOS device?  There's an R package called 'itunesr' that allows for free extracts of user ratings and reviews.  Combining that with some sentiment analysis, you can examine any trends for apps of interest.

First, you'll need to identify an app of interest.  I use the <br>Starbucks</br> app quite a bit and have read numerous articles about it's success.  So, let's see if 'itunesr' tells a smiliar story.  

Let's install the needed programs.

```{r}
# Set CRAN Mirror
options(repos = c(CRAN = "http://cran.rstudio.com"))

# Set time zone
options(tz="America/New_York")

# install itunesr directly from CRAN:
install.packages("itunesr")
# the development version from GitHub:
# install.packages("devtools")
# devtools::install_github("amrrs/itunesr")
library(itunesr)
```

Now, let's find the app's ID.  To do this, you must search for it in the apple online store and you'll find the number in the URL.  <a href="https://apps.apple.com/us/app/starbucks/id331177714">Here</a> is Starbucks' page and you'll note that its ID is 331177714.

```{r}
# Starbucks page:  https://apps.apple.com/us/app/starbucks/id331177714
# Latest (Page 1) Starbucks Reviews for the Country: US
sbux_reviews <- getReviews(331177714,'us',1)
head(sbux_reviews,2)
```

This is a lot of data!  Let's start off by looking at the numerical ratings/scores that users have given the app.

```{r}
# Ratings Trend
install.packages("highcharter")
# install.packages("lubridate")
# install.packages("dplyr")
library(highcharter)
library(dplyr)
library(lubridate)
dt <- sbux_reviews
dt$Date <- as.Date(dt$Date)
dt$Rating <- as.numeric(dt$Rating)    
dt <- dt %>% select(Date,Rating) %>% group_by(Date) %>% summarise(Rating = round(mean(Rating),2))
```
```
# Plot
highchart() %>%   hc_add_series_times_values(dt$Date,dt$Rating, name = 'Average Rating')
```
<img src = "/post/2019-09-16-starbucks-ios-app-analysis_files/avg_rating.png">

We see that the rating is generally pretty high.  But, this doesn't take advantage of all the text users have provided.  To measure that, let's use Bing's defined sentiments to illustrate how postivie/negative the feedback is.

```{r}
# Sentiment Analysis 
install.packages("sentimentr")
library(sentimentr)
library(tidytext)
library(stringr)
# identify and define sentiments/words
get_sentiments("bing")
```

Next, we'll want to combine the sentiment scores with the words used in the reviews.

```{r}
# pull in sbux info
reviews_only <- as.character(sbux_reviews$Review)
head(reviews_only,2)
sentiment_scores <- reviews_only %>% sentiment_by(by=NULL)
```

```
# plot findings
highchart() %>% hc_xAxis(sentiment_scores$element_id) %>% hc_add_series(sentiment_scores$ave_sentiment, name = 'Reviews Sentiment Scores')
```

<img src = "/post/2019-09-16-starbucks-ios-app-analysis_files/sent_score.png">

Now we have two visualizations that give us information on our app of interest: Starbucks.  This can easily be duplicated for any others... and it's absolutely FREE.

[THE END]